---
title: "Visualisation"
author: "Philipp"
date: "1/11/2019"
output:
  bookdown::html_document2:
    toc: true
    theme: darkly
---

# Introduction to ggplot2

Learning how to use the ggplot2 package to visualize sequence data. 

```{r Setup, echo=FALSE}
library(ggplot2)
```

```{r Read&setup_data, echo=FALSE, include=FALSE}
variantData <- read.delim("http://genomedata.org/gen-viz-workshop/intro_to_ggplot2/ggplot2ExampleData.tsv")
head(variantData)
variantData[1:5, c(1:9,16,18)]
variantData$tumor_COV <- variantData$tumor_ref_count + variantData$tumor_var_count
```

## basics

The following results in the same graphic, however, be aware in figure p2 that you have to follow this order, or at least the `geom_point()` argument with the `aes()` in the end, otherwise the plot kinda does not work.

```{r plot1&2, echo=TRUE, fig.cap="Figure 1&2: Comparison of two ways to create the same plot"}
p1 <- ggplot(data = variantData, aes(x = tumor_VAF, y = tumor_COV))
p1 <- p1 + geom_point()
p1 <- p1 + theme_classic()
p1 <- p1 + ggtitle('p1')
p1

p2 <- ggplot()+
  theme_classic()+
  ggtitle('p2')+
  geom_point(data = variantData, aes(x = tumor_VAF, y = tumor_COV))
p2
```

Now we are going to spead out the data a little more...

```{r plot3, echo=FALSE, fig.cap="Figure 3: Jitter"}
p3 <- ggplot()+
  geom_point(data = variantData, aes(x = chromosome_name, y = tumor_VAF), position = 'jitter')
p3
```

Limting the axis of the plots

```{r plot4, echo=FALSE, fig.cap="Figure 4: limiting the axis"}
p4 <- p1 + scale_y_continuous(limits = c(0,500))
p4
```

```{r plot5, echo=FALSE, fig.cap="Figure 5: transforming the actual values in the data frame"}
p5 <- ggplot()+
  theme_classic()+
  geom_point(data = variantData, aes(x = tumor_VAF, y = log2(tumor_COV)), size = .7, shape = 3)
p5
```

```{r plot6, echo=FALSE, fig.cap="Figure 6: transform the axis using ggplot"}
p6 <- p1 + scale_y_continuous(trans = 'log2') + ylab('log2(tumor_COV)')
p6
```

## Applying different aesthetics

```{r plot7, echo=FALSE, fig.cap="Figure 7: what happens when we try to add color within the aesthetics?"}
p7 <- ggplot()+
  geom_point(data = variantData, aes(x = tumor_VAF, y = tumor_COV, color = 'blue'))+
  scale_y_continuous(trans = 'log2')
p7
```

```{r plot8, echo=FALSE, fig.cap="Figure 8: and what happens when we try to add color within the geom"}
p8 <- ggplot()+
  geom_point(data = variantData, aes(x = tumor_VAF, y = tumor_COV), color = 'blue')+
  scale_y_continuous(trans = 'log2')
p8
```

```{r plot9, echo=FALSE, fig.cap="Figure 9: get a density curve of tumor vafs"}
p9 <- ggplot()+
  geom_density(data = variantData, aes(x = tumor_VAF, color = dataset))
p9
```

`adjust =` in the `geom_density` allows for smoothing of the density function.

```{r plot10, echo=FALSE, fig.cap="Figure 10: let's add a bit more detail"}
p10 <- ggplot()+
  geom_density(data = variantData, aes(x = tumor_VAF, fill = dataset), alpha = .5, color = 'black', adjust = .5)
p10
```

```{r plot11, echo=FALSE, fig.cap="Figure 11: and let's change the colors some more"}
p11 <- p10+
  scale_fill_brewer(palette = 'Set1')
p11
```

```{r plot12, echo=FALSE, fig.cap="Figure 12: putting different colors..."}
p12 <- ggplot()+
  geom_density(data = variantData, aes(x = tumor_VAF, fill = dataset, color = dataset))+
  scale_fill_brewer(palette = 'Greys')+
  theme_classic()
p12
```

## Faceting

